---
title: "lab 1"
author: "Claudia Carugati"
format: pdf
editor: visual
---

## Quarto

## Running Code

```{r}
library(tidyverse)
library(viridis)
```

```{r}
glimpse(midwest)
```

```{r}
view(midwest)
```

1.  Making a histogram to visualize the population density of counties

```{r}
ggplot(midwest, aes(x = popdensity))+
  geom_histogram(binwidth = 10000, fill = "light blue")+
labs(title = "population density in different counties", x = "population density", y = "frequency")
```

The distribution has a right skewed tail.

There seems to be a few counties with an extremely high population density at around 60000 and 90000. These are outliers, they are far away from most of the data and outside of the curve.

2.  Create a scatterplot of the percentage of people with a college degree versus percentage below poverty

    ```{r}
    ggplot(midwest, aes(x=percollege, y=percbelowpoverty, color=state))+
      geom_point()+
      labs(title="% of people in college vs % of people below poverty line", x= "% of people in college", y= "% of people below poverty line") + scale_colour_viridis_d()
    ```

<!-- -->

3.  Describe what you observe in the plot from the previous exercise. In your description, include similarities and differences in the patterns across states.

4.  exercise title

    ```{r}
    ggplot(midwest, aes(x=percollege, y=percbelowpoverty, color=state))+
      geom_point(alpha = .5)+
      facet_wrap(~state) +
      geom_smooth(se = FALSE, color = "black")+
      labs(title="% of people in college vs % of people below poverty line", x= "% of people in college", y= "% of people below poverty line") + scale_colour_viridis_d()
    ```

    Which plot do you prefer - this plot or the plot in Ex 2? Briefly explain your choice.

5.  doing something? - add title

    ```{r}
    ggplot(midwest, aes(x=area, y=state))+
      geom_boxplot()
    ```

-   Describe what you observe from the plot.

-   Which state has the single largest county? How do you know based on the plot?

6.  plotting something ? - add title

```{r}
midwest <- midwest |>
  mutate(metro = if_else(inmetro == 1, "Yes", "No"))
```


```{r}
ggplot(midwest, aes(x=state, fill= metro))+
  geom_bar()
```
7. recreating a plot 

```{r}
    ggplot(midwest, aes(x=percollege, y=popdensity, color=percbelowpoverty))+
      geom_point(size=2, alpha=0.5)+
       facet_wrap(~state) +
      labs(title="Do people with college degrees tend to live in denser areas?", x= "% college educated", y= "population density (person/unit area)", fill= "% below \n poeverty line")
    ```

